---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r knitr-setup, include = FALSE, results = "asis", comment = ""}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Setting up and using [`ricu`](https://cran.r-project.org/package=ricu) to access the publicly available ICU database [AmsterdamUMCdb](https://github.com/AmsterdamUMC/AmsterdamUMCdb) of [Amsterdam UMC](https://www.amsterdamumc.nl).

---

**With release of `ricu` version 0.2.0, this repository has become obsolete and only serves for illustration purposes on how to set up a new data source with `ricu`.**

---

In order to use the AmsterdamUMCdb as `ricu`-external dataset `aumc_ext`, some configuration is necessary to be set up before loading `ricu`. A configuration file `data-sources.json` provides the necessary info on the involved tables and `concept-dict.json` is required for loading `ricu` clinical concepts from the new dataset. Finally some dataset-specific implementations of certain S3 generic functions exported from `ricu` are required (see `r/ricu.R`) as might be some callback functions used in concept specification (see `r/callback.R`).

## Configuring ricu

Several environment variables can be set for `ricu` to facilitate the integration of `aumc_ext`:

* `RICU_DATA_PATH`: (optionally) point this to a directory containing a folder `aumc_ext` which holds the AmsterdamUMCdb data
* `RICU_CONFIG_PATH`: point this to the `./config` directory of this project
* `RICU_SRC_LOAD`: Comma separated list of data sources to automatically load when attaching `ricu` (e.g. `mimic,mimic_demo,aumc_ext`)

For now, we leave `RICU_DATA_PATH` at default and set the other environment variables accordingly before loading `ricu`

```{r, config-setup}
sources <- c("mimic_demo", "eicu_demo", "aumc", "aumc_ext")

Sys.setenv(RICU_SRC_LOAD = paste(sources, collapse = ","),
           RICU_CONFIG_PATH = "config")

library(ricu)

for (file in file.path("r", c("ricu.R", "callback.R"))) {
  source(file)
}
```

As `aumc` is already fully supported by `ricu`, all configuration in this repo has been renamed to set up the AmsterdamUMCdb data as external dataset `aumc_ext`.

## Set-up

In order to download and set up the data as `aumc_ext`, but using `ricu` functionality for downloading and importing `aumc`, run:

```{r, data-setup, eval = FALSE}
download_src("aumc", src_data_dir("aumc_ext"))
import_src("aumc", src_data_dir("aumc_ext"))
attach_src("aumc_ext")
```

This mimics manual download and conversion to `.fst` files (as one might do for setting up a new data source manually) and mimimizes the amount of configuration information during set-up (compare `aumc_ext` of `data-sources.json` provided by this repo with `aumc` of the `ricu` provided `data-sources.json`.)

## Load data

Upon successful set-up, data can be loaded as:

```{r, data-load}
aumc_ext$processitems

load_ts(aumc_ext$processitems, itemid == 9159, id_var = "patientid")
load_ts(aumc_ext$processitems, itemid == 9159, id_var = "admissionid")

gluc <- concept("gluc", unit = "mmol/l",
  item("aumc_ext", "numericitems", "itemid", list(c(9947L, 6833L, 9557L)))
)

load_concepts(gluc, id_type = "patient", verbose = FALSE)
load_concepts(gluc, id_type = "icustay", verbose = FALSE)

concept_availability(concepts = c("glu", "alb", "weight"))

load_concepts(c("glu", "alb"), "aumc_ext", verbose = FALSE)
load_concepts(c("alb", "weight"), "aumc_ext", verbose = FALSE)
```
